\dppTitle

### 2014-06-02 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
dpp ms ready for resubmission.

\change{
Recently, Oaks et al.\ \cite{Oaks2012,Oaks2014reply} found via simulation that
\msb will often strongly support models with a small number of divergence
events shared among taxa, even when divergences were random over broad
timescales.}
They suggested this behavior was due to the combination of uniform priors on
parameters causing small marginal likelihoods of richer models and the
U-shaped prior on divergence models.
Hickerson et al.\ \cite{Hickerson2013} suggested the problem was caused by
sampling error, and proposed as a solution an approximate-Bayesian model
averaging approach that samples over empirically informed uniform priors.
However, Oaks et al.\ \cite{Oaks2014reply} evaluated the approach proposed by
Hickerson et al.\ \cite{Hickerson2013} using simulations and found that it did
not mitigate the method's propensity to incorrectly infer clustered
divergences, and often preferred priors that excluded the true values of the
model's parameters.
\change{
Here, I describe a new method that successfully mitigates spurious inference of
co-divergence while avoiding negative side effects of empirically informed
uniform priors.}

@@

In the original descriptions of the \msb method \cite{Hickerson2006,Huang2011},
this re-ordering is justified by the fact that the expected value of $\pi_b$ is
unrelated to sample size \popSampleSize{1}{} and \popSampleSize{2}{} and thus
exchangeable among pairs.
\change{
This is incorrect for two reasons.}
First, the entire vector of summary statistics \alignmentSS{}{} for each pair
of populations is re-ordered across pairs, which implies that the justification
for re-ordering $\pi_b$ applies to all the statistics within each
\alignmentSS{}{}.
\change{
However, the expectations for statistics that estimate gross diversity (e.g.,
$\pi$ and $\theta_W$) are not independent of sample size for structured
populations (e.g., the divergent pairs of populations modeled by \msb), and
other statistics are not independent of sample size in general (e.g.,
$SD(\pi-\theta_W)$).
Second, and more importantly, having the same expectation does not ensure
random variables are exchangeable.
Rather, for variables to be exchangeable their marginal distributions must be
the same (i.e., they must be identically distributed).
None of the summary statistics used by \msb, including $\pi_b$, are identically
distributed when there is any variation among taxa or loci in the
(1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
(2) length of alignments,
(3) numbers of loci (\nloci{}),
(4) \hky model parameters (\hkyModel{}{}),
(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
or
(6) ploidy multipliers (\ploidyScalar{}{}).
Whenever such variation is present (i.e., nearly all empirical applications),
the taxon-specific summary statistics \alignmentSS{}{} are not exchangeable,
and the reshuffling of the summary statistic vectors is not mathematically
valid.

@@

\section{Discussion}
My results demonstrate that using alternative priors on parameters and
divergence models improved the behavior of the \msb model.
In the new implementation, model-choice estimation is more accurate and shows
greater robustness to model violations
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The original model is very sensitive to violations and, when present,
strongly over-estimates the probability of one-divergence event shared
across all taxa
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
\change{
When more appropriate priors are used for divergence-time and demographic
parameters, and either a Dirichlet-process or uniform prior applied across
divergence models, the model is less sensitive to violations, and, when
violations do cause bias, the method is conservative and tends to underestimate
the probability of models with temporally clustered divergences
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).}

>>>>

\change{
Recently, Oaks et al.\ \cite{Oaks2012,Oaks2014reply} found via simulation that
\msb will often strongly support models with a small number of divergence
events shared among taxa, even when divergences were random over broad
timescales.}
They suggested this behavior was due to the combination of uniform priors on
parameters causing small marginal likelihoods of richer models and the
U-shaped prior on divergence models.
Hickerson et al.\ \cite{Hickerson2013} suggested the problem was caused by
sampling error, and proposed as a solution an approximate-Bayesian model
averaging approach that samples over empirically informed uniform priors.
However, Oaks et al.\ \cite{Oaks2014reply} evaluated the approach proposed by
Hickerson et al.\ \cite{Hickerson2013} using simulations and found that it did
not mitigate the method's propensity to incorrectly infer clustered
divergences, and often preferred priors that excluded the true values of the
model's parameters.
\change{
Here, I describe a new approach that successfully mitigates spurious inference
of co-divergence while avoiding negative side effects of empirically informed
uniform priors.}

@@

In the original descriptions of the \msb method \cite{Hickerson2006,Huang2011},
this re-ordering is justified by the fact that the expected value of $\pi_b$ is
unrelated to sample size \popSampleSize{1}{} and \popSampleSize{2}{} and thus
exchangeable among pairs.
\change{
This is incorrect for two reasons.}
First, the entire vector of summary statistics \alignmentSS{}{} for each pair
of populations is re-ordered across pairs, which implies that the justification
for re-ordering $\pi_b$ applies to all the statistics within each
\alignmentSS{}{}.
\change{
However, the expectations for statistics that estimate gross diversity (e.g.,
$\pi$ and $\theta_W$) are not independent of sample size for structured
populations (e.g., the divergent pairs of populations modeled by \msb), and
other statistics are not independent of sample size in general (e.g.,
$SD(\pi-\theta_W)$).
Second, and more importantly, having the same expectation does not ensure
random variables are exchangeable.
Rather, for variables to be exchangeable their marginal distributions must be
the same (i.e., they must be identically distributed).
\emph{None} of the summary statistics used by \msb, including $\pi_b$, have
this property when there is any variation among taxa or loci in the
(1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
(2) length of alignments,
(3) numbers of loci (\nloci{}),
(4) \hky model parameters (\hkyModel{}{}),
(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
or
(6) ploidy multipliers (\ploidyScalar{}{}).
Whenever such variation is present (i.e., nearly all empirical applications),
the taxon-specific summary statistics \alignmentSS{}{} are not exchangeable,
and the reshuffling of the summary statistic vectors is not mathematically
valid.

@@

\section{Discussion}
My results demonstrate that using alternative priors on parameters and
divergence models improved the behavior of the \msb model.
In the new implementation, model-choice estimation is more accurate and shows
greater robustness to model violations
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The original model is very sensitive to violations and, when present,
strongly over-estimates the probability of one-divergence event shared
across all taxa
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
\change{
When more appropriate priors are used for divergence-time and demographic
parameters, and either a Dirichlet-process or uniform prior applied across
divergence models, the model is less sensitive to violations, and, when
violations do cause bias, the method tends to underestimate
the probability of models with temporally clustered divergences
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
Given that clustered models are often of particular interest to biogeographers,
this behavior of the new method can be considered conservative.
}
